---
title: "Roulette Simulation"
author: "Fisher Ankney"
date: "4/24/2018"
output: html_document
---

<br> 

### Introduction 

In this post, all about roulette! Now everyone knows the odds of roulette, bet on black, almost 50% chance, bet thirds, almsot a 33% chance, bet a number, no chance! Disecting the probability is easy, we know that the house will always win, sample size increases. Lets make a story out of it. three people at a roulette table with their own styles of playing. Phabian always bets $10 on black. Tammy always bets \$ on the middle third. Latigo always bets \$10 on his favorite number, 23. 

How do they fair against each other after 10 spins? 100 spins? 1000 spins? 10,000 spins?!?
number of wins, amount won, highest high, lowest low, that kind of thing. how many numbers go by before winnning on average for the three players. how long is their longest dry streak? how many wins in a row? 

Going to finish up with creating a simple roulette function, allowing you to place your bet on a number or category, how much you bet on that spin. returns the number, if you win or lose, and your cumulative winnings. 

also, how old are they for sitting through 10,000 spins of the roulette wheel?
- 10,000 spins at 45 spins an hour 9.25 days of continuous play 

roulette(your_number, your_bet, new_game = F)
returns win or loss, cummulative sum of winnings / losings 

i.e.

roulette(6, 100)
-> ** 18 ** 
-> You Lose! 
-> Total Winnings: -$1500

```{r}
library('tidyverse')
```

Creating the data
```{r}
roulette_full <- sample(-1:36, 10000, replace=T)

roulette_ten <- roulette_full[1:10]
roulette_hundred <- roulette_full[1:100]
roulette_thousand <- roulette_full[1:1000]
```



```{r}
# Initialize scenario

#make each of these an empty vector like examine vector, then say something along the 
# lines of if win, then gambler_a_cash[[i]] = gambler_a_cash[[i-1]] + 20
gambler_a_cash <- 0 
gambler_a_wins <- 0 

gambler_b_cash <- 0 
gambler_b_wins <- 0 

gambler_c_cash <- 0
gambler_c_wins <- 0 



```
















```{r}

# Define Bets
black <- c(2, 4, 7, 8, 10, 11, 13, 15, 17, 20, 22, 24, 26, 28, 29, 31, 33, 35)
third_dozen <- c(25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36)
lucky_23 <- 23


# Initialize Variables
ga_cash <- vector("double", length(roulette_ten))
ga_total_cash <- 0 
ga_wins <- vector("double", length(roulette_ten))
ga_total_wins <- 0 

gb_cash <- vector("double", length(roulette_ten))
gb_total_cash <- 0 
gb_wins <- vector("double", length(roulette_ten))
gb_total_wins <- 0 

gc_cash <- vector("double", length(roulette_ten))
gc_total_cash <- 0 
gc_wins <- vector("double", length(roulette_ten))
gc_total_wins <- 0 

for (i in seq_along(roulette_full)) {
  # All Gamblers Bet
  ga_total_cash = ga_total_cash - 10
  gb_total_cash = gb_total_cash - 10
  gc_total_cash = gc_total_cash - 10
  # Gambler A Wins
  if (roulette_full[[i]] %in% black) {
    ga_total_cash = ga_total_cash + 20 
    ga_total_wins = ga_total_wins + 1 
  }
  # Gambler B Wins
  if (roulette_full[[i]] %in% third_dozen) {
    gb_total_cash = gb_total_cash + 30 
    gb_total_wins = gb_total_wins + 1
  }
  # Gambler C Wins
  if (roulette_full[[i]] == lucky_23) {
    gc_total_cash = gc_total_cash + 360
    gc_total_wins = gc_total_wins + 1
  }
  # End of Round
  ga_cash[[i]] <- ga_total_cash
  ga_wins[[i]] <- ga_total_wins
  gb_cash[[i]] <- gb_total_cash
  gb_wins[[i]] <- gb_total_wins
  gc_cash[[i]] <- gc_total_cash
  gc_wins[[i]] <- gc_total_wins
}

```
  
    gambler_a_cash[[i]] = gambler_a_cash
    gambler_a_wins[[i]] = gambler_a_wins[[i-1]] + 1
  }

if num in roulette 2, 4, 7, 8, 10, 11, 13, 15, 17, 20, 22, 24, 26, 28, 29, 31, 33, 35
then phabian_win + 1 && phabian_money + $ 20 


for-loop snippet
```
output <- vector("double", ncol(df))  # 1. output
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output

```


for num in roulette 
phabian_money -$10
tammy_money - $10
latigo_money - $10

if num in roulette 2, 4, 7, 8, 10, 11, 13, 15, 17, 20, 22, 24, 26, 28, 29, 31, 33, 35
then phabian_win + 1 && phabian_money + $ 20 

else if num in roulette = 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36 
then tammy_win + 1 && tammy_money + $30

else if num in roulette = 23 
then Latigo_win + 1 && latigo_money + $360 

for each roulette spin, keep track of number wins and $ winnings 

I want to plot it all at the end 






