---
title: "Roulette Simulation"
author: "Fisher Ankney"
date: "4/24/2018"
output: html_document
---

<br> 

### Introduction 

Recenty I rewatched the movie *Support your Local Gunfighter*, a classic western from 1971. Several times throughout the movie, the protagonist Latigo places his signature 23-black bet on a game of roulettete. While Latigo went all in on his lucky number, I began to dream up a simple roulette simulation in R.

I've devised a short study to evaluate three roulette strategies over a period of time in order to practice for-loop and if-then statements in R. In this simulation, Gambler A always bets $10 on black, which has a 47.37% chance of winning. Gambler B always bets \$10 on the third dozen numbers, which has a 32.43% chane of winning. Gambler C, Latigo, Always bets \$10 on 23 black, which has a 2.7% change of winning. 

By creating a sample of 10,000 roulette spins and running a simulation with these three gambling strategies, I will evaluate how each gambler performs at differing intervals. The number of wins and total earnings of each gambler, how these trends change over time, and the occurances and duration of *dry streaks* are all of interest in this simulation. 

I'll finish up the post by creating a simple roulette function in R that will satisfy all of your gambling desires. But I' be willing to bet that those desires will be long gone after we evaluate the long term earnings associated with playing roulette. 

<br> 

### Library

```{r, message=F}
library('tidyverse')
```

<br> 

### Simulation

A seperate script created to simulate 10,000 roulette spins - input/roulette_sim.R 
```{r, eval=F}
library('tidyverse')

roulette <- as_tibble(sample(-1:36, 10000, replace=T))
write_csv(roulette, "roulette.csv")

```

<br> 

Loading the spins - 
```{r, eval=F}
roulette <- read_csv("roulette.csv")
roulette <- as.integer(roulette)
str(roulette)
```

<br> 

```{r}
# The Gamblers Bets
black <- c(2, 4, 7, 8, 10, 11, 13, 15, 17, 20, 22, 24, 26, 28, 29, 31, 33, 35)
third_dozen <- c(25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36)
lucky_23 <- 23

# Initialize gambler A
ga_cash <- vector("double", length(roulette))
ga_total_cash <- 0 
ga_wins <- vector("double", length(roulette))
ga_total_wins <- 0 
# Initialize gambler B
gb_cash <- vector("double", length(roulette))
gb_total_cash <- 0 
gb_wins <- vector("double", length(roulette))
gb_total_wins <- 0 
# Initialize gambler C
gc_cash <- vector("double", length(roulette))
gc_total_cash <- 0 
gc_wins <- vector("double", length(roulette))
gc_total_wins <- 0 

for (i in seq_along(roulette)) {
# All gamblers bet $10
  ga_total_cash = ga_total_cash - 10
  gb_total_cash = gb_total_cash - 10
  gc_total_cash = gc_total_cash - 10
# If gambler A wins, +$20 and +1 win
  if (roulette[[i]] %in% black) {
    ga_total_cash = ga_total_cash + 20 
    ga_total_wins = ga_total_wins + 1 
  }
# If gambler B wins, +$30 and +1 win
  if (roulette[[i]] %in% third_dozen) {
    gb_total_cash = gb_total_cash + 30 
    gb_total_wins = gb_total_wins + 1
  }
# If gambler C wins, +$360 and +1 win
  if (roulette[[i]] == lucky_23) {
    gc_total_cash = gc_total_cash + 360
    gc_total_wins = gc_total_wins + 1
  }
# End of round - save totals 
  ga_cash[[i]] <- ga_total_cash
  ga_wins[[i]] <- ga_total_wins
  gb_cash[[i]] <- gb_total_cash
  gb_wins[[i]] <- gb_total_wins
  gc_cash[[i]] <- gc_total_cash
  gc_wins[[i]] <- gc_total_wins
}
```

<br> 

### Evaluation 

Plot roulette dataset to check for patterns

```{r}
plot(roulette)
```

<br> 

```{r}
ga <- "Gambler A"
gb <- "Gambler B"
gc <- "Gambler C"

results <- tibble(spin_num = 1:10000, ga, ga_cash, ga_wins, gb, gb_cash, gb_wins, gc, gc_cash, gc_wins)
results_100 <- head(results, n=100)
results_1000 <- head(results, n=1000)
```

<br> 

```{r}
ggplot(results_100) + 
  geom_point(aes(x = spin_num, y = ga_wins, color = ga), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y = gb_wins, color = gb), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y  = gc_wins, color = gc), alpha = 0.8) + 
  theme_minimal() +
  theme(legend.title=element_blank()) + 
  scale_colour_manual(values = c("black", "#E0080B", "#016D29")) + 
    labs(title = "100 Spin Results - Wins", 
       y = "Number of Wins",
       x = "Number of Spins"
       )
```

<br> 

```{r}
ggplot(results_100) + 
  geom_point(aes(x = spin_num, y = ga_cash, color = ga), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y = gb_cash, color = gb), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y = gc_cash, color = gc), alpha = 0.8) + 
  theme_minimal() +
  theme(legend.title=element_blank()) + 
  scale_colour_manual(values = c("black", "#E0080B", "#016D29")) + 
    labs(title = "100 Spin Results - Earnings", 
       y = "Earnings ($)",
       x = "Number of Spins"
       )
```

<br>

```{r}
ggplot(results_1000) + 
  geom_point(aes(x = spin_num, y = ga_cash, color = ga), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y = gb_cash, color = gb), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y  = gc_cash, color = gc), alpha = 0.8) + 
  theme_minimal() +
  theme(legend.title=element_blank()) + 
  scale_colour_manual(values = c("black", "#E0080B", "#016D29")) + 
    labs(title = "1000 Spin Results - Earnings", 
       y = "Earnings ($)",
       x = "Number of Spins"
       )
```

<br>

```{r}
ggplot(results) + 
  geom_point(aes(x = spin_num, y = ga_wins, color = ga), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y = gb_wins, color = gb), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y  = gc_wins, color = gc), alpha = 0.8) + 
  theme_minimal() +
  theme(legend.title=element_blank()) + 
  scale_colour_manual(values = c("black", "#E0080B", "#016D29")) + 
    labs(title = "10,000 Spin Results - Wins", 
       y = "Number of Wins",
       x = "Number of Spins"
       )
```

<br>

```{r}
ggplot(results) + 
  geom_point(aes(x = spin_num, y = ga_cash, color = ga), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y = gb_cash, color = gb), alpha = 0.8) + 
  geom_point(aes(x = spin_num, y  = gc_cash, color = gc), alpha = 0.8) + 
  theme_minimal() +
  theme(legend.title=element_blank()) + 
  scale_colour_manual(values = c("black", "#E0080B", "#016D29")) + 
    labs(title = "10,000 Spin Results - Earnings", 
       y = "Earnings ($)",
       x = "Number of Spins"
       )
```

<br>

trendline for 10,000, average lose per spin for each gambler, max / min, box plot for dry streak each player 

```{r}
ga_total_wins / 10000 * 100
gb_total_wins / 10000 * 100
gc_total_wins / 10000 * 100
```

<br> 

actual probability of each win - 
half bet = 46.37%
thirds bet = 31.58%
single number bet = 2.63%

Average loss - 
```{r}
ga_total_cash / 10000
gb_total_cash / 10000
gc_total_cash / 10000
```

<br> 

```{r}
min(ga_cash)
min(gb_cash)
min(gc_cash)

```

<br> 

```{r, error=T}
ga_wins_lead <- lead(ga_wins)
gb_wins_lead <- lead(gb_wins)
gc_wins_lead <- lead(gc_wins) 

ga_wins_binary <- ga_wins_lead - ga_wins
ga_wins_binary[10000] = 0
gb_wins_binary <- gb_wins_lead - gb_wins
gb_wins_binary[10000] = 0 
gc_wins_binary <- gc_wins_lead - gc_wins
gc_wins_binary[10000] = 0

dry_count_a <- 0 
dry_count_save_a  <- vector("double", length(ga_wins_binary))
dry_count_b <- 0
dry_count_save_b  <- vector("double", length(gb_wins_binary))
dry_count_c <- 0
dry_count_save_c  <- vector("double", length(gc_wins_binary))


 
for (i in seq_along(ga_wins_binary)) {
  if (ga_wins_binary[[i]] == FALSE) {
    dry_count_a = dry_count_a + 1
  }
  if (ga_wins_binary[[i]] == TRUE) {
    dry_count_save_a[[i]] = dry_count_a
    dry_count_a = 0
  }
}

dry_count_save_a
dry_tibble_a <- tibble("dry_streak" = dry_count_save_a)

ggplot(dry_tibble_a) + 
  geom_bar(aes(dry_streak))
  
for (i in seq_along(gb_wins_binary)) {
  if (gb_wins_binary[[i]] == FALSE) {
    dry_count_b = dry_count_b + 1
  }
  if (gb_wins_binary[[i]] == TRUE) {
    dry_count_save_b[[i]] = dry_count_b
    dry_count_b = 0
  } 
}

dry_count_save_b
dry_tibble_b <- tibble("dry_streak" = dry_count_save_b)

ggplot(dry_tibble_b) + 
  geom_bar(aes(dry_streak))

for (i in seq_along(gc_wins_binary)) {
  if (gc_wins_binary[[i]] == FALSE) {
    dry_count_c = dry_count_c + 1
  }
  if (gc_wins_binary[[i]] == TRUE) {
    dry_count_save_c[[i]] = dry_count_c
    dry_count_c = 0
  } 
}

dry_count_save_c
dry_tibble_c <- tibble("dry_streak" = dry_count_save_c)

ggplot(dry_tibble_c) + 
  geom_bar(aes(dry_streak))

```


create a roulette function

```{r, eval = F}
mini_roulette <- function(bet, num) {

  print(str_c("You've bet $", bet, " on ", num))
  result = sample(0:12,1)
  print(str_c("The ball has landed on ", result))
  
  if ( num == result ) {
    print(str_c("You've Won $", bet * 36, "!"))
    total_winnings = total_winnings + bet * 36
  } else {
    print ("You Lose!")
    total_winnings = total_winnings - bet
  }
  print(str_c("Your total winnings are ", total_winnings, " dollars"))
}

mini_roulette(100, 0)
mini_roulette(100, 1)
mini_roulette(100, 2)
```
