---
title: "hands_on_r_2"
author: "Fisher Ankney"
date: "9/29/2018"
output: html_document
---

Chapter 7 - Programs

randomly generate some symbols!

diamonds - DD 
seven - 7 
Bar Bar Bar - BBB
Bar Bar - BB 
Bar - B
Cherry - C
Zero - 0 

```{r}
get_symbols <- function() {
  wheel <- c('DD', '7', 'BBB', 'BB', 'B', 'C', '0')
  symbols <- sample(
         wheel,
         size = 3,
         replace = T,
         prob = c(0.03, 0.03, 0.06, 0.1, 0.25, 0.01, 0.52)
        )
  print(symbols)
}

```

Test it 
```{r}
get_symbols()
```


we're going to create a function to score each roll based on a set of rules

two types of sutasks, parallel and sequencial 

sequential steps - 
# step 1 
do this
# step 2 
do that
# step 3 
print conclusions

parallel cases - 
# evaluate conditions
# if way 1 
do way 1 
# if way 2 
do way 2
# if way 3 
do way 3
# print results
print

The slot machine program will look something like this - 

slot_machine() - 
get symbols
display symbols

three branches - 
- 3 of a kind, score type 1 
- all bars, score type 2 
- other method, score type 3
- and so on

adjust for diamonds (multiplier)

print score 

## If statements 
```
if (this) {
  that
}
```

fun little function 
```{r}
a <- 3.14
trunc(a)
a - trunc(a)
```




test whether the scores are three of a kind
```{r}
symbols <- c("7", "7", "7")

if (symbols[[1]] == symbols[[2]] & symbols[[2]] == symbols[[3]]) {
  print("true")
}

# another way to do it 
if (length(unique(symbols)) == 1) {
  print('its true')
}
```


now add this test to the slot machine script 

```{r, eval = F}
same <- length(unique(symbols)) == 1 

if (same) {
  prize <- # look up prize $ 
} else if ( # case 2: all bars ) {
  prize <- # $5
} else {
  # count cherries
  prize <- # calculate a prize
}
 # calculate diamonds, use as multiplier 
```



now we test if all bars 
```{r}
symbols <- c('B', 'BB', 'BBB')


all(symbols %in% c('B', 'BB', 'BBB'))
```

great, insert this into the program 
```{r, eval=F, error=T}
same <- length(unique(symbols)) == 1 
bars <- all(symbols %in% c('B', 'BB', 'BBB'))
if (same) {
  prize <- # look up prize $ 
} else if (bars) {
  prize <- # $5
} else {
  # count cherries
  prize <- # calculate a prize
}
 # calculate diamonds, use as multiplier 
```

now to assign prize values 

don't use a long and tedious if then tree, use a lookup table
```{r}
payouts <- c('DD' = 100, '7' = 80, 'BBB' = 40, 'BB' = 25, 'B' = 10, 'C' = 10, '0' = 0)

payouts

payouts[['DD']]

payouts[[symbols[1]]]
```

add it to the slot machine game
```{r, eval=F, error=T}
same <- length(unique(symbols)) == 1 
bars <- all(symbols %in% c('B', 'BB', 'BBB'))
payouts <- c('DD' = 100, '7' = 80, 'BBB' = 40, 'BB' = 25, 'B' = 10, 'C' = 10, '0' = 0)


if (same) {
  prize <- payouts[[symbols[1]]] 
} else if (bars) {
  prize <- # $5
} else {
  # count cherries
  prize <- # calculate a prize
}
 # calculate diamonds, use as multiplier 
```


now we go tot he next task, assign $5 prize to bars
```{r, eval=F, error=T}
same <- length(unique(symbols)) == 1 
bars <- all(symbols %in% c('B', 'BB', 'BBB'))
payouts <- c('DD' = 100, '7' = 80, 'BBB' = 40, 'BB' = 25, 'B' = 10, 'C' = 10, '0' = 0)


if (same) {
  prize <- payouts[[symbols[1]]] 
} else if (bars) {
  prize <- 5
} else {
  # count cherries
  prize <- # calculate a prize
}
 # calculate diamonds, use as multiplier 
```


count cherries and diamonds
```{r}
symbols <- c("C", "0", "C")

sum(symbols == "C")

symbols <- c("D", "0", "0")

sum(symbols == "D")
```

Add these to the program
```{r, eval = F, error=T}
same <- length(unique(symbols)) == 1 
bars <- all(symbols %in% c('B', 'BB', 'BBB'))
payouts <- c('DD' = 100, '7' = 80, 'BBB' = 40, 'BB' = 25, 'B' = 10, 'C' = 10, '0' = 0)


if (same) {
  prize <- payouts[[symbols[1]]] 
} else if (bars) {
  prize <- 5
} else {
  cherries <- sum(symbols == "C")
  prize <- # calculate a prize
}
  diamonds <- sum(symbols == "D")
  prize <- # calculate prize
```


now calculate cherries prize 
```{r, eval = F}
symbols <- c("C", "C", "0")

# create a lookup table
prize <- c(0,2,5)[cherries + 1]

# must +1 to index properly, because zero cherries is first index
```

Put into program 
```{r}
same <- length(unique(symbols)) == 1 
bars <- all(symbols %in% c('B', 'BB', 'BBB'))
payouts <- c('DD' = 100, '7' = 80, 'BBB' = 40, 'BB' = 25, 'B' = 10, 'C' = 10, '0' = 0)


if (same) {
  prize <- payouts[[symbols[1]]] 
} else if (bars) {
  prize <- 5
} else {
  cherries <- sum(symbols == "C")
  prize <- c(0,2,5)[cherries + 1]
}
  diamonds <- sum(symbols == "D")
  prize <- # calculate prize
```

diamond multiplier 
```{r}
prize * 2 ^ diamonds
```

insert code and comment 
```{r}
# check if all symbols are the same
same <- length(unique(symbols)) == 1
# check if all symbols are variety of bar
bars <- all(symbols %in% c('B', 'BB', 'BBB'))

# if all symbols are the same
if (same) {
# assign payout value in 'lookup table'
  payouts <- c('DD' = 100, '7' = 80, 'BBB' = 40, 'BB' = 25, 'B' = 10, 'C' = 10, '0' = 0)
  prize <- payouts[[symbols[1]]] 
# if all symbols are bars
} else if (bars) {
# prize is $5
  prize <- 5
# else check for cherries
} else {
# count cherries
  cherries <- sum(symbols == "C")
# assign prize based on cherry
  prize <- c(0,2,5)[cherries + 1]
}
# now check if any diamonds are present
diamonds <- sum(symbols == "D")
# adjust prize as necissary 
prize <- prize * 2 ^ diamonds
```

package it into a function





```{r}
get_symbols <- function() {
  wheel <- c('DD', '7', 'BBB', 'BB', 'B', 'C', '0')
  sample(
         wheel,
         size = 3,
         replace = T,
         prob = c(0.03, 0.03, 0.06, 0.1, 0.25, 0.01, 0.52)
        )
}

```




```{r}
score <- function(symbols) {
  
  # check if all symbols are the same
  same <- length(unique(symbols)) == 1
  # check if all symbols are variety of bar
  bars <- all(symbols %in% c('B', 'BB', 'BBB'))
  
  # if all symbols are the same
  if (same) {
  # assign payout value in 'lookup table'
    payouts <- c('DD' = 100, '7' = 80, 'BBB' = 40, 'BB' = 25, 'B' = 10, 'C' = 10, '0' = 0)
    prize <- payouts[[symbols[1]]] 
  # if all symbols are bars
  } else if (all(bars)) {
  # prize is $5
    prize <- 5
  # else check for cherries
  } else {
  # count cherries
    cherries <- sum(symbols == "C")
  # assign prize based on cherry
    prize <- c(0,2,5)[cherries + 1]
  }
  
  # now check if any diamonds are present
  diamonds <- sum(symbols == "D")
  # adjust prize as necissary 
  prize <- prize * 2 ^ diamonds

}
```

now put them together
```{r}
play <- function() {
  symbols <- get_symbols()
  print(symbols)
  score(symbols)
  
}
```

now test
```{r}
play()
```

## The S3 System

s3 refers to a class system built into R. governs how R handels objects of different classes. 

built around three components - attributes, generic functions, and methods 

### attributes 
attributes are peices of extra information that are given a name and appended to an object. 

dont affect the value, but stick to it as metadata
```{r}
attributes()
```

add attributes with function attr
```{r}
# create var one_play
one_play <- play() 

# show var one_play
one_play

# show attributes of var one_play
attributes(one_play)

# add symbols attribute to one_play
attr(one_play, "symbols") <- c("B","0","B")

# show attributes one_play
attributes(one_play)
```


Modify play to return a prize that contains the symbols associated with it as an attribute named symbols. remove the redundant call to print(symbols)
```{r}
play <- function() {
  symbols <- get_symbols()
  prize <- score(symbols)
  attr(prize, 'symbols') <- symbols
  prize
}
```

lets see if it works 
```{r}
play()
```

ugly, but we can fix that 
```{r}
two_play <- play()
two_play
```

```{r}
slot_display <- function(prize){
  
# extra symbol 
symbols <- attr(prize, "symbols")

# collapse symbols into single string 
symbols <- paste(symbols, collapse = " ")

# combine symbols with prize as a regex
string <- paste(symbols, prize, sep = '\n$')

#display regex in consol without quotes
cat(string)

}
```

test it
```{r}
slot_display(play())
```

this method of cleaning the output requires you to manually intervene in your R session. There is a function to automatically clean it up, and that's print(). 

## generic functions

```{r}
class(one_play) <- "slots"

print.slots <- function(x, ...) {
  cat("I'm using print.slots method")
}

print(one_play)
```

excellent, so lets print slots the way we want to print them

```{r}
print.slots <- function(x, ...) {
  slot_display(x)
}
```

now try it out
```{r}
one_play
```

now to get every single play() to have attribute slot
```{r}
play <- function() {
  symbols <- get_symbols()
  structure(score(symbols), symbols = symbols, class= 'slots')
}

class(play())
```

```{r}
play()
```


