---
title: "Hands on R Programming Resource Review Notes"
author: "Fisher Ankney"
date: "9/27/2018"
output: html_document
---

Why coding is better than a GUI - Reproducability, Automation, Communication 

Project 1 - elementary 

Project 2 - Playing Cards 
- chapter 3 - R Objects 

## atomic vector

atomic vector in R
- grouping some values together using c(), or just a single value

```{r}
vector <- c(1,2,3)
is.vector(vector)
```

```{r}
number <- 1
is.vector(number)
```

Types of atomic vectors - 
doubles - regular number, default num 
integer - num followed by uppercase "L"
character - string of text 
logical - T/F 
complex and raw - don't worry about these

names - attribute to give an atomic vector 
```{r}
die <- c(1,2,3,4,5,6)
names(die) <- c("one", "two", "three","four", "five", "six")
die
```

to remove names, set to null
```{r}
names(die) <- NULL
die
```

## atomic array

manipulate an atomic array by giving it a dim (dimension)
```{r}
dim(die) <- c(2,3) # 2 row, 3 col
die
dim(die) <- c(3,2)
die
```

## matrices 

```{r}
m <- matrix(die, nrow = 2)
m

m <- matrix(die, nrow = 2, byrow = TRUE) #fill by row 
m
```

## more info on your object

```{r}
attributes(m)
```


```{r}
class(m)
```
 

```{r}
str(m)
```

## Dates and times 

```{r}
now <- Sys.time() 
now
```

## Factors 
way of storing categorical information 

```{r}
# create a vactor with factor()
gender <- factor(c("male", "female", "male", "female"))

# check type of object 'gender'
typeof(gender)

# check attributes 'gender'
attributes(gender)

# see underlying structure of 'gender'
unclass(gender)

# print 'gender'
gender

# coercing data structure
char_gender <- as.character(gender)
char_gender
```

## coercion 

everything coerces to character if they're present
logical will go numeric if it must

```{r}
# logical to numeric
sum(c(T,F,T))

# numeric to chr
char <- c(1,"2")
typeof(char)

log_to_char <- c(TRUE, "TRUE")
typeof(log_to_char)
```

## Lists 
```{r}
# first entry coerced, second double, third char, fourth logical
list <- list(c(1:10, "yeah"), 1:10, "nope", TRUE, list("you found me",2))
list

# a list is a list
typeof(list)

# but elements of the list don't have to be 
typeof(list[[1]])

# unless you want them to 
typeof(list[[5]])

# access this list of a list by 
list[[5]][[1]]
```


but lists are ugly, 

## dataframes
most useful storage structure for data anlysis 
cant combine different lengths 

```{r}
library('tidyverse')

read_

```


```{r}
deck <- as_data_frame(read_csv(url("https://gist.githubusercontent.com/garrettgman/9629323/raw/ee5dfc039fd581cb467cc69c226ea2524913c3d8/deck.csv")))

deck

# save the dataset just incase you go offline, or the source is removed
write_csv(deck, "deck_data.csv")
```


## chapter 4 - R notation

selecting elements of a data frame 
```{r}
# positive integers
deck[1,1] # first element
deck[1,] # first row
deck[,1] # first column 
```

Negative integers, everything but these rows
```{r}
deck[,-1] # everything but first column
```

can also make logical
```{r}
deck[TRUE,c(FALSE,TRUE,FALSE)] # all row, only col 2 
```


function to deal the deck 
```{r}
deck2 <- deck

deal <- function(cards) {
  cards[1, ]
  }

deal(deck)
```

random deal tho
```{r}
random <- sample(1:52)
random

deck[random,]

deal <- function(cards) {
  random <- sample(1:52)
  cards[random, ]
}

deal(deck)
```

## Modifying Values - chapter 5 

changing values in place 
```{r}
values <- c(0,0,0,0,0)
values[1]

values[1] <- 1000
values

```

great way to add variables to dataset is in place
```{r}
deck$new <- 1:52
deck
```


or to destroy them
```{r}
deck$new <- NULL
deck
```

== not = to evaluate
```{r}
a = 7
a
a == 7
```

```{r}
sum(deck$face == "jack")
```

change the values of all the jacks
```{r}
deck$value[deck$face == "jack"] <- 1000
deck
```

logical subsetting is one thing R does super well
a key to writing vectorized code, chapter 10 

in hearts every ard has a value zero except hearts, which have 1 
```{r}
deck$value[deck$suit == "hearts"] <- 1 
deck$value[deck$suit != "hearts"] <- 0 
deck$value
```


boolean operators
```{r}
# & | xor ! any(x,y,z) all(x,y,z)
# and or only-1-or not any all
```


give the queen of spades value 13
```{r}
queen_of_spades <- deck$face == "queen" & deck$suit == "spades"
deck$value[queen_of_spades] <- 13 
deck[queen_of_spades, ]
```

ready to play hearts


extra exercises 
```{r}
w <- c(-1,0,1)
x <- c(5,15)
y <- "February"
z <- c("monday",'tuesday', 'wednesday')

# is w positive
w > 0 
# is x greater than 10 and less than 20 
x > 10 & x < 20 
# is object y the word february
y == "February"
# is every value in z a day of the week
all(z %in% c('monday', 'tuesday', 'wednesday', 'thursday', 'friday', 'saturday', 'sunday'))
```


new deck 
```{r}
deck <- as_data_frame(read_csv(url("https://gist.githubusercontent.com/garrettgman/9629323/raw/ee5dfc039fd581cb467cc69c226ea2524913c3d8/deck.csv")))

deck
```

Missing Information 
many functions come with na.rm
is.na is a useful function 
NA's are contageous 



